---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SBIFM

<!-- badges: start -->
<!-- badges: end -->

The goal of SBIFM is sparse modelling of high-dimensional covariance matrices using Bayesian latent factor models. The method proposed by [Bhattachrya et. al.](https://doi.org/10.1093/biomet/asr013) has been used to develop an efficient Gibbs sampler that scales well as data dimensionality increases. For automatically truncating the infinite loading matrix through selection of the number of important factors, we use the much more efficient truncation criteria proposed by [Schiavon et. al.]( https://doi.org/10.1002/sta4.298). 


## Installation

You can install the development version of SBIFM from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("RivuROX913/SBIFM")
```

## Example

This is a basic example which shows applications of different functions:

```{r example generation}
library(SBIFM)

## Generate synthetic data
set.seed(2143)
data = generateData(n = 200, p = 30, k = 10, rep = 5)
tail(data)
```

The function `generateData()` generates a sparse loading matrix with given parameters and returns factor model data with that loading matrix along with other true parameters.

```{r example analysis}
## Analysis
out = GibbsCov(data, 10000, 3000, 7, 1e-4)
```

The function `GibsCov()` runs the MCMC algorithm to estimate the covariance matrix and number of factors needed. If the true Covariance matrix is included in the data, it also returns the error in estimation:

```{r convergence}
## Error in covariance estimation
out$Error

## Convergence rate
plot(out$post.factor[1:(data$p+10)], type = "l", xlab = "Iteration", ylab = "Number of factors", main = "Evolution of factors")
```

Above plot shows the average factors in the model across iterations for multiple replications of the data. As we can see, the convergence rate of the factors is almost linear.

# Real data

The function `imputeData()` was made to help users in handling real datasets. To show a brief example we consider the `bfi` dataset included in the `psych` package that contains 25 personality items taken from the International Item Pool. For more details regarding the data, [click here](https://pmagunia.com/dataset/r-dataset-package-psych-bfi).

`imputeData()` function can make a list suitable as an input to the function `GibbsCov()`. It can also impute any missing data if necessary. Following plots visualize the amount of missing data before and after applying the function:

```{r real}
# Real data and visualisation
library(psych)
library(dlookr, quietly = TRUE)
library(visdat)

## Dataset
data = bfi[, 1:25]
head(data)

#diagnose for missing value
diagnose(data)
vis_miss(data, sort_miss = FALSE)

#imputation
datalist = imputeData(data)

##completed data
vis_miss(datalist$data, sort_miss = FALSE)
```

Applying `GibbsCov()` on above imputed dataset shows that the 25 personality traits actually depends on only 6 factors:

```{r Real data analysis}
out = GibbsCov(datalist, 20000, 5000, 3, 1e-4)
out$Factor
```

We can also find one of the estimates of the loading matrix, even though it is not identifiable.

```{r}
head(out$Lambda)
```

